### 2019-06-11 initial/notes.tex
stash

\section{Weak Tidal Friction, changing $\eta$}

@@

First, we note that the two integrals that I earlier called $\Delta_m, \Delta_c$
(garbage notation sorry) can be evaluated for the new weak tidal friction
integrals as follows:
\begin{align}
    \Delta_M &= \int\limits_0^{2\pi}\epsilon\p*{1 - \mu^2}
            \p*{\frac{2\Omega}{s} - \mu} \;\mathrm{d}\phi,\nonumber\\
        &\approx 2\pi \epsilon \frac{2\Omega}{s},\\
    \Delta_c &= \oint \epsilon \p*{1 - \mu^2}\p*{\frac{2\Omega}{s} - \mu}
            \;\mathrm{d}\phi,\nonumber\\
        &= 2\int\limits_0^{2\pi}
            \epsilon\sqrt{2\eta \sin I\p*{1 - \cos \phi}}\;\mathrm{d}\phi
            ,\nonumber\\
        &= 16\epsilon\sqrt{\eta \sin I}.
\end{align}
Recalling that $\eta = s_c/s$, we arrive at
\begin{equation}
    P_{hop} = \frac{4\sqrt{ss_c\sin I}}{\pi \Omega}.
\end{equation}

@@

First, let's examine our $t_{\star}$ plot (TODO add). We observe that there is a
roughly linear slope $\rd{t_{\star}}{(\Delta H)} \sim \frac{600}{s_0/2s_c}$
where $-s_0/2s_c$ is the minimum value of $H$ on the domain and so is the
maximum value for $\Delta H = H_4 - H_0$. Note that
\begin{align*}
    H(\mu, \phi; s) &= -\frac{s}{s_c}\frac{\mu^2}{2}
        + \mu \cos I - \sin I \sqrt{1 - \mu^2}\cos \phi, &
    H_4(s) &= -\sin I + \frac{s_c}{2s}\cos^2 I.
\end{align*}
While $s$ is actually a dynamical variable, it is a parameter to the 1D
Hamiltonian. Plugging these expressions ino our above expression for $\rd{K}{t}
= -\rd{(\Delta H)}{t}$, we obtain
\begin{align}
    \rd{(\Delta H)}{t} &= -\dot{\mu}^{(1)}\dot{\phi}^{(0)} - \dot{s}^{(1)}
            \rd{(\Delta H)}{s},\\
    -\rd{(\Delta H)}{(\epsilon t)} &= (1 - \mu^2)\p*{\frac{2\Omega}{s} - \mu}
            \dot{\phi}^{(0)} + 2\Omega\p*{1 + \frac{s}{2\Omega}(1 + \mu^2)}
            \s*{\frac{\mu^2}{2s_c} + \sin I + \frac{s_c}{2s^2}\cos^2 I}.
\end{align}
Now, if we evaluate far from the separatrix and for large $s$, similar to most
of our initial conditions, we approximate $\dot{\phi} \approx \frac{s_c}{s}$ and
we obtain
\begin{equation}
    \rd{(\Delta H)}{t} \approx \epsilon
        \frac{s\mu^2}{s_c}\p*{1 - \mu^2 + \frac{1 + \mu^2}{2}} \sim
        -\epsilon\frac{s\mu^2}{s_c}.
\end{equation}
Comparing to our plot, we would predict that this derivative should be close to
$\frac{s/s_c}{1200}$ (assuming $s \sim s_0$ initially), so it seems like we at
least have the scaling correct.

@@

>>>>

\section{Parametric Separatrix Hopping: Only changing $\eta$}

Though this section is located before the weak tidal friction section, we
actually return to this second toy problem to get some practice on the Henrard
paper, to confirm the accuracy of our simulations compared to analytical
predictions, and to get some practice doing these new integrals. Consider the
dissipation-free Cassini state system with a small but constant $\pd{\eta}{t} =
\epsilon$. For clarity, the Hamiltonian is
\begin{equation}
    H = -\frac{\mu^2}{2} + \eta
        \p*{\mu \cos I - \sin I \sqrt{1 - \mu^2} \theta \cos \phi}.
\end{equation}

For an arbitrary initial condition outside of the separatrix, we want: (i) to
know when it will encounter the separatrix, and (ii) what its fate will be. We
will begin with (ii) since it is the more fun problem. We know that this boils
down to the Melnikov-like formula
\begin{equation}
    B_i = \int_{\mathcal{C}_i} \dot{\eta}\pd{\Delta H}{\eta}
        \;\mathrm{d}t.
\end{equation}
$B_i$ is the change in $\Delta H = H_4 - H$ over contour $\mathcal{C}_i$.

\section{Weak Tidal Friction, changing $\eta$ and $\theta$}

@@

First, we note that the two integrals that I earlier called $\Delta_m, \Delta_c$
(garbage notation sorry) can be evaluated for the new weak tidal friction
integrals as follows:
\begin{align}
    \Delta_M &= \int\limits_0^{2\pi}\epsilon\p*{1 - \mu^2}
            \p*{\frac{2\Omega}{s} - \mu} \;\mathrm{d}\phi,\nonumber\\
        &\sim 2\pi \epsilon \p*{\frac{2\Omega}{s}
            - \sqrt{2\frac{s_c}{s}\sin I}},\\
    \Delta_c &= \oint \epsilon \p*{1 - \mu^2}\p*{\frac{2\Omega}{s} - \mu}
            \;\mathrm{d}\phi,\nonumber\\
        &= 2\int\limits_0^{2\pi}
            \epsilon\sqrt{2\eta \sin I\p*{1 - \cos \phi}}\;\mathrm{d}\phi
            ,\nonumber\\
        &= 16\epsilon\sqrt{\eta \sin I}.
\end{align}
Recalling that $\eta = s_c/s$, we arrive at
\begin{equation}
    P_{hop} = \frac{4\sqrt{ss_c\sin I}}{\pi \Omega}
        \p*{1 + \sqrt{\frac{s_cs\sin I}{2}}}.\label{eq:p_hop_sqrt_s}
\end{equation}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.7\textwidth]{../initial/1_weaktide/4_cross0_20.png}
    \caption{$t_\star, s_\star$ for $s_c = 0.2$. Red line is a by-eye linear
    fit.}\label{fig:4_cross0_20}
\end{figure}

First, let's examine our $t_{\star}$ plot e.g.\ \autoref{fig:4_cross0_20}. We
observe that there is a roughly linear slope $\rd{t_{\star}}{(\Delta H)} \sim
\frac{600}{s_0/2s_c}$ where $-s_0/2s_c$ is the minimum value of $H$ on the
domain and so is the maximum value for $\Delta H = H_4 - H_0$. Note that
\begin{align*}
    H(\mu, \phi; s) &= -\frac{s}{s_c}\frac{\mu^2}{2}
        + \mu \cos I - \sin I \sqrt{1 - \mu^2}\cos \phi, &
    H_4(s) &= -\sin I + \frac{s_c}{2s}\cos^2 I.
\end{align*}
While $s$ is actually a dynamical variable, it is a parameter to the 1D
Hamiltonian. Plugging these expressions into our above expression for $\rd{K}{t}
= -\rd{(\Delta H)}{t}$, we obtain
\begin{align}
    \rd{(\Delta H)}{t} &= -\dot{\mu}^{(1)}\dot{\phi}^{(0)} - \dot{s}^{(1)}
            \pd{(\Delta H)}{s},\\
    -\rd{(\Delta H)}{(\epsilon t)} &= (1 - \mu^2)\p*{\frac{2\Omega}{s} - \mu}
            \dot{\phi}^{(0)} + 2\Omega\p*{1 + \frac{s}{2\Omega}(1 + \mu^2)}
            \s*{\frac{\mu^2}{2s_c} - \frac{s_c}{2s^2}\cos^2 I}.
\end{align}
Note that this expression is encouraging since both $\dot{\phi}^{(0)},
\pd{\p*{\Delta H}}{s}$ vanish near CS4 (where $\mu_4 \approx \frac{s_c}{s}\cos
I$), which corresponds to separatrix approach slowing down (also consistent w/
our plots). Now, if we evaluate far from the separatrix and for large $s$,
similar to most of our initial conditions, we approximate $\dot{\phi} \approx
\frac{s_c}{s}$ and we obtain
\begin{align}
    \rd{(\Delta H)}{(\epsilon t)} &\approx
            \p*{1 - \mu^2}\p*{\frac{2\Omega}{s} - \mu}\frac{s}{s_c}\mu
                + s\p*{1 + \mu^2}\p*{\frac{\mu^2}{2s_c}},\nonumber\\
        &\approx -\frac{s\mu^2}{2s_c} \p*{2(1 - \mu^2) - 1 + \mu^2}
            ,\nonumber\\
        &= -\frac{s\mu^2}{2s_c}\p*{3 - \mu^2}.
\end{align}
The linear fit by eye on the plot corresponds to a slope of $-\rd{(\Delta H)}{t}
\approx \frac{s_0/s_c}{1200}$ which is of the right order $-\epsilon
\frac{3s\mu^2}{2s_c}$, so it seems like we at least have the scaling correct.
The intercept offset is given by the dynamics near $\Delta H$ small, or
sufficiently small that $\dot{\phi}^{(0)} \approx \frac{s}{s_c}$ is no longer
accurate (where phase space begins to be distorted by the separatrix); we will
analyze this separately later/never.

\subsection{$P_{hop}$ More Carefully}

@@

It bears noting that the formula \autoref{eq:p_hop_sqrt_s} only holds under the
following assumptions: (i) the Melnikov splitting is stronger than the adiabatic
invariant effect described by Henrard, and (ii) $\frac{2\Omega}{s} \gtrsim \mu$
along the separatrix. Condition (ii) is expected to hold in the weak perturber
regime $s_c$ small, and indeed it seems to give the correct prediction, as seen
in \autoref{fig:4_cross_hist0_05} (a few other trials also seem to suggest
this).
\begin{figure}[t]
    \centering
    \includegraphics[width=0.7\textwidth]{../initial/1_weaktide/4_cross_hist0_05.png}
    \caption{Histogram of $P_{hop}(s_\star)$ with prediction of
    \autoref{eq:p_hop_sqrt_s} overlaid. Anal indicates just the prefactor, while
    corrected analytic denotes including the estimated corrective term. We see
    that the true curve lies somewhere in between.}\label{fig:4_cross_hist0_05}
\end{figure}

Can we analytically show that the Melnikov splitting dominates, and that in
particular the seeming small excess of $P_{hop}$ is predictable?
